### 2018-01-30 Publication_bias_analyses.Rmd
CogSci final prep; publication bias markdown file

---
title: "MP Publication Bias"
author: "originally Christina Bergmann, modified by Katie Von Holzen"
output: 
  pdf_document:
    toc: true
---
```{r Setup, echo = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, tidy = TRUE, message = FALSE, error = TRUE)

### Load libraries
library(metafor)
library(pwr)
library(ggplot2)
library(grid)
library(gridExtra)

```

#Preparation

Read in data and tidy up dataset

```{r ReadIn,  echo = FALSE}
#Get the data
source("scripts/Meta_Analysis.R")

# change name of mispronunciation sensitivity object
db_mpe <- dat
rm(dat)

```

# Funnel Plot Asymmetry
Use ranktest() to check for funnel plot asymmetry. All three tests show evidence of significant asymmetry.

```{r FunnelPlotAsymmetry, echo = FALSE}

# Publication bias can become visible in funnel plot asymmetry. 
# Metafor comes with several options to check this, I chose ranktest and regtest

# this is the model of the mispronunciation effect, without age
ranktest(rma_MPeffect)

# correct object identification
ranktest(rma_correct)

# mispronounced object identification
ranktest(rma_MP)

```

# Export data for p-curve

This step creates a text file that can be used in the p-curve app:
http://www.p-curve.com/app4/

These files can be found in "p_curve_app/"

for correct: p_curve_co.txt
for mispronounced: p_curve_mp.txt

Just copy and paste the .txt onto the website and voila!

NOTE: This will always append lines to the existing txt file, check for double writing!

```{r PcurveExport, echo = FALSE}

textfile_co = "p_curve_app/p_curve_co.txt"

for(line in 1:length(db_ET_correct$n_1)){
  if(!is.na(db_ET_correct[line,]$t)){
    newline = paste("t(", db_ET_correct[line,]$n_1-1, ")=",db_ET_correct[line,]$t, sep = "")
    write(newline, file = textfile_co, append = TRUE)
    #cat("\n")
    #write(newline, file = textfile, append = TRUE)    
  }
}

textfile_mp = "p_curve_app/p_curve_mp.txt"

for(line in 1:length(db_ET_MP$n_1)){
  if(!is.na(db_ET_MP[line,]$t)){
    newline = paste("t(", db_ET_MP[line,]$n_1-1, ")=",db_ET_MP[line,]$t, sep = "")
    write(newline, file = textfile_mp, append = TRUE)
    #cat("\n")
    #write(newline, file = textfile, append = TRUE)    
  }
}
```

# Calculate Power

# What is the power of the studies in the dataset?

```{r Power_exist, echo = FALSE}
d = rma.mv(g_calc, g_var_calc, data = db_mpe, random = ~ 1 | short_cite)$b[,1]

# estimate from mispronunciation sensitivity
mp_est <- coef(summary(rma_MPeffect))[2,]$estimate

pwr.t.test(n = median(db_mpe$n_1, na.rm=TRUE), d = mp_est, sig.level = .05, type = "paired", alternative = "two.sided")

```

# What is the average power to detect a mispronunciation sensitivity effect (to be compared with p-curve estimate)
```{r Power_plan, echo = FALSE}

# estimate from mispronunciation sensitivity
mp_est <- coef(summary(rma_MPeffect))[2,]$estimate

# one-sided t-test
pwr.t.test(d = mp_est, sig.level = .05, type = "paired", power = .80, alternative = "greater")

```

# Draw Funnel Plots

### Plots from Sakaluk, 2016
### https://sakaluk.wordpress.com/2016/02/16/7-make-it-pretty-plots-for-meta-analysis/

## Funnel Plot for Correct Object Identification

```{r FunnelPrep, echo = FALSE}

#Themes and plot
apatheme=theme_bw()+
  theme(#panel.grid.major=element_blank(),
    #panel.grid.minor=element_blank(),
    #panel.border=element_blank(),
    axis.line=element_line(),
    text=element_text(family='Times', size=25),
    legend.position='none')

```

## Funnel Plot for Correct Object Identification

```{r FunnelCorrect, echo = FALSE}

#Rename a bunch of things for ease
dat_co = db_ET_correct
#dat_co=db_ET_correct
dat_co <-
  dat_co %>% 
  rename(cite = short_cite,
         yi = g_calc,
         vi = g_var_calc)

#Reorder bibliographic info based on value of g (yi), so effect sizes can be plotted in descending order

dat_co <-
  dat_co %>% 
  select(cite, expt_num, same_infant, mean_age_1, yi, vi) %>% 
  mutate(study_ref = paste(cite, expt_num, same_infant, sep=',')) %>% 
  arrange(desc(yi))

#Get standard errors from variances
dat_co$se = sqrt(dat_co$vi)

#Calculate 95% CI values
dat_co$lowerci = (-1.96*dat_co$se)+dat_co$yi
dat_co$upperci = (1.96*dat_co$se)+dat_co$yi

#Store the meta-analytic estimate and its standard error from whatever model you run
#summary(rma_correct)
estimate_co = coef(summary(rma_correct))$estimate
se_co = coef(summary(rma_correct))$se

#Store a vector of values that spans the range from 0
#to the max value of impression (standard error) in your dataset.
#Make the increment (the final value) small enough (I choose 0.001)
#to ensure your whole range of data is captured
se.seq_co=seq(0, max(dat_co$se), 0.001)

#Now, compute vectors of the lower-limit and upper limit values for
#the 95% CI region, using the range of SE that you generated in the previous step, 
#and the stored value of your meta-analytic estimate.
ll95_co = estimate_co-(1.96*se.seq_co)
ul95_co = estimate_co+(1.96*se.seq_co)

#You can do this for a 99% CI region too
ll99_co = estimate_co-(3.29*se.seq_co)
ul99_co = estimate_co+(3.29*se.seq_co)

#And finally, do the same thing except now calculating the confidence interval
#for your meta-analytic estimate based on the stored value of its standard error
meanll95_co = estimate_co-(1.96*se_co)
meanul95_co = estimate_co+(1.96*se_co)

#Now, smash all of those calculated values into one data frame (called 'dfCI').
#You might get a warning about '...row names were found from a short variable...'
#You can ignore it.
dfCI_co = data.frame(ll95_co, ul95_co, ll99_co,
                  ul99_co, se.seq_co, estimate_co, meanll95_co, meanul95_co)

#Now we can actually make the funnel plot.
#Using your original data-frame, map standard error to your x-axis (for now) and Zr to your y-axis
fp_co = ggplot(aes(x = se, y = yi), data = dat_co) +
  #Regression line for the FP asymmetry
  geom_smooth(aes(x = se, y = yi), method = "lm", colour = "darkgrey", alpha = .5, se = FALSE, data = dat_co) +
  #Add your data-points to the scatterplot
  geom_point(size = 2.5, colour="black") +
  #
  #Give the x- and y- axes informative labels
  xlab('Standard Error') + ylab('Hedge\'s g')+
  # make sure both plots have same scale
  # still ylim because we haven't flipped yet
  #coord_cartesian(ylim = c(-1, 2.5))+
  #Now using the 'dfCI' data-frame we created, plot dotted lines corresponding
  #to the lower and upper limits of your 95% CI region
  #And dashed lines corresponding to your 99% CI region
  #Add lines corresponding to 0 and estimate
  geom_line(aes(x = se.seq_co, y = 0), linetype = 'solid', data = dfCI_co) +
  geom_line(aes(x = se.seq_co, y = estimate_co), linetype = 'dashed', data = dfCI_co) +
  geom_line(aes(x = se.seq_co, y = ll95_co), linetype = 'dotted', data = dfCI_co) +
  geom_line(aes(x = se.seq_co, y = ul95_co), linetype = 'dotted', data = dfCI_co) +
  #  geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
  #  geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
  #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
  #geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI) +
  #geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI) +
  #Reverse the x-axis ordering (se) so that the tip of the funnel will appear
  #at the top of the figure once we swap the x- and y-axes...
  scale_x_reverse()+
  #Specify the range and interval for the tick-marks of the y-axis (Zr);
  #Choose values that work for you based on your data
  #scale_y_continuous(breaks=seq(-.45,0.8,0.25))+
  #And now we flip the axes so that SE is on y- and Zr is on x-
  coord_flip()+
  #Finally, apply my APA-format theme (see code at end of post).
  #You could, alternatively, specify theme_bw() instead.
  apatheme

#Call the pretty funnel plot
fp_co +   ggtitle("Correct")
ggsave("figures/FunnelPlot_correct.pdf")
```

## Funnel Plot for Mispronunciation Object Identification

```{r FunnelMisp, echo = FALSE}

  #Rename a bunch of things for ease
dat_mp = db_ET_MP  
#dat_mp=db_ET_MP
  dat_mp <-
    dat_mp %>% 
    rename(cite = short_cite,
           yi = g_calc,
           vi = g_var_calc)

  #Reorder bibliographic info based on value of g (yi), so effect sizes can be plotted in descending order

  dat_mp <-
    dat_mp %>% 
    select(cite, expt_num, same_infant, mean_age_1, yi, vi) %>% 
    mutate(study_ref = paste(cite, expt_num, same_infant, sep=',')) %>% 
    arrange(desc(yi))

  #Get standard errors from variances
  dat_mp$se = sqrt(dat_mp$vi)

  #Calculate 95% CI values
  dat_mp$lowerci = (-1.96*dat_mp$se)+dat_mp$yi
  dat_mp$upperci = (1.96*dat_mp$se)+dat_mp$yi

  #Store the meta-analytic estimate and its standard error from whatever model you run (substitute your own values)
  summary(rma_MP)
  estimate_mp = coef(summary(rma_MP))$estimate
  se_mp = coef(summary(rma_MP))$se

  #Store a vector of values that spans the range from 0
  #to the max value of impression (standard error) in your dataset.
  #Make the increment (the final value) small enough (I choose 0.001)
  #to ensure your whole range of data is captured
  se.seq_mp=seq(0, max(dat_mp$se), 0.001)

  #Now, compute vectors of the lower-limit and upper limit values for
  #the 95% CI region, using the range of SE that you generated in the previous step, 
  #and the stored value of your meta-analytic estimate.
  ll95_mp = estimate_mp-(1.96*se.seq_mp)
  ul95_mp = estimate_mp+(1.96*se.seq_mp)

  #You can do this for a 99% CI region too
  ll99_mp = estimate_mp-(3.29*se.seq_mp)
  ul99_mp = estimate_mp+(3.29*se.seq_mp)

  #And finally, do the same thing except now calculating the confidence interval
  #for your meta-analytic estimate based on the stored value of its standard error
  meanll95_mp = estimate_mp-(1.96*se_mp)
  meanul95_mp = estimate_mp+(1.96*se_mp)

  #Now, smash all of those calculated values into one data frame (called 'dfCI').
  #You might get a warning about '...row names were found from a short variable...'
  #You can ignore it.
  dfCI_mp = data.frame(ll95_mp, ul95_mp, ll99_mp,
                       ul99_mp, se.seq_mp, estimate_mp, meanll95_mp, meanul95_mp)

  #Now we can actually make the funnel plot.
  #Using your original data-frame, map standard error to your x-axis (for now) and Zr to your y-axis
  fp_mp = ggplot(aes(x = se, y = yi), data = dat_mp) +
    #Regression line for the FP asymmetry
    geom_smooth(aes(x = se, y = yi), method = "lm", colour = "darkgrey", alpha = .5, se = FALSE, data = dat_mp) +
    #Add your data-points to the scatterplot
    geom_point(size = 2.5, colour="black") +
    #
  #Give the x- and y- axes informative labels
  xlab('Standard Error') + ylab('Hedge\'s g')+
    # make sure both plots have same scale
    # still ylim because we haven't flipped yet
    #coord_cartesian(ylim = c(-1, 2.5))+
    #Now using the 'dfCI' data-frame we created, plot dotted lines corresponding
    #to the lower and upper limits of your 95% CI region
    #And dashed lines corresponding to your 99% CI region
    #Add lines corresponding to 0 and estimate
    geom_line(aes(x = se.seq_mp, y = 0), linetype = 'solid', data = dfCI_mp) +
    geom_line(aes(x = se.seq_mp, y = estimate_mp), linetype = 'dashed', data = dfCI_mp) +
    geom_line(aes(x = se.seq_mp, y = ll95_mp), linetype = 'dotted', data = dfCI_mp) +
    geom_line(aes(x = se.seq_mp, y = ul95_mp), linetype = 'dotted', data = dfCI_mp) +
    #  geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
    #  geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
    #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
    #geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI) +
    #geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI) +
    #Reverse the x-axis ordering (se) so that the tip of the funnel will appear
    #at the top of the figure once we swap the x- and y-axes...
    scale_x_reverse()+
    #Specify the range and interval for the tick-marks of the y-axis (Zr);
    #Choose values that work for you based on your data
    #scale_y_mpntinuous(breaks=seq(-.45,0.8,0.25))+
    #And now we flip the axes so that SE is on y- and Zr is on x-
    coord_flip()+
    #Finally, apply my APA-format theme (see code at end of post).
    #You could, alternatively, specify theme_bw() instead.
    apatheme

  #Call the pretty funnel plot
  fp_mp + ggtitle("Mispronunciation")
  ggsave("figures/FunnelPlot_misp.pdf")

```

## Combine both funnel plots for figure

```{r FunnelCombo, echo = FALSE}

  jpeg(filename = "figures/Funnel_Plots_Object_Identification.jpg", 
       width = 600, height = 400, units = "px")

  p <-   grid.arrange(fp_co, fp_mp, nrow = 1)
                      #top = textGrob("Object Identification", gp=gpar(fontsize=25)))
                      #top = "Object Identification")

  dev.off()

    grid.draw(p)

```