### 2018-02-05 CogSci_Analyses.Rmd
Created CogSci markdown files for Analysis and Publication Bias
Markdown files have been uploaded to OSF CogSci Subproject.

---
title: "Mispronunciation MetaAnalysis for CogSci"
author: "Christina Bergmann and Katie Von Holzen"
output: 
  pdf_document:
    toc: true
---
```{r Setup, echo = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, tidy = TRUE, message = FALSE, error = TRUE)

### Load libraries
library(tidyverse)
library(metafor)
library(meta)
library(pwr)
library(knitr)
library(wesanderson)

```

#Preparation

Read in data and tidy up dataset

```{r ReadIn,  echo = FALSE}
#Get the data
source("scripts/calculateES.R")

```

```{r Preprocess, echo = FALSE}

db_ET <- db_ET %>%
  mutate(age.C = (mean_age_1-mean(mean_age_1, na.rm=TRUE))/30.44) %>%
  filter(mean_age_months < 31)

db_ET$collapse <- paste(db_ET$study_ID, db_ET$expt_num, db_ET$same_infant, sep = "_")

#Split into correct and MP database

db_ET_correct <- db_ET[db_ET$is_correct=="1",]
db_ET_MP <- db_ET[db_ET$is_mp=="1",]

#collapse over nonindependent MP rows for a *general* MP effect

collapse_rows <- function(db){
  db$collapse <- paste(db$study_ID, db$expt_num, db$same_infant, sep = "_")

  for(independent in unique(db$collapse)){
    if(length(db[db$collapse==independent])>1){
      sub = db[db$collapse==independent, ]
      sub$d_calc <- median(sub$d_calc)
      sub$d_var_calc <- median(sub$d_var_calc)
      sub$g_calc <- median(sub$g_calc)
      sub$g_var_calc <- median(sub$g_var_calc)
      sub$corr_imputed <- median(sub$corr_imputed)
      db <- db[!(db$collapse==independent),]
      db <- rbind(db, sub[1,])
    }
  }
  return(db)
}

#db_ET_correct = collapse_rows(db_ET_correct)

#db_ET_MP = collapse_rows(db_ET_MP)

#remove outliers, for now we have none, though
db_ET_MP$nooutlier = ifelse(db_ET_MP$g_calc > mean(db_ET_MP$g_calc, na.rm = TRUE) + 3*sd(db_ET_MP$g_calc, na.rm = TRUE) 
                         | db_ET_MP$g_calc < mean(db_ET_MP$g_calc, na.rm = TRUE) - 3*sd(db_ET_MP$g_calc, na.rm = TRUE),FALSE, TRUE)
db_ET_MP = db_ET_MP[db_ET_MP$nooutlier,]

db_ET_correct$nooutlier = ifelse(db_ET_correct$g_calc > mean(db_ET_correct$g_calc, na.rm = TRUE) + 3*sd(db_ET_correct$g_calc, na.rm = TRUE) 
                         | db_ET_correct$g_calc < mean(db_ET_correct$g_calc, na.rm = TRUE) - 3*sd(db_ET_correct$g_calc, na.rm = TRUE),FALSE, TRUE)
db_ET_correct = db_ET_correct[db_ET_correct$nooutlier,]

# make sure that both correct and mispronounced conditions are considered in descriptives

db_ET_correct$condition <- 1
db_ET_MP$condition <- 0

dat <- bind_rows(db_ET_correct, db_ET_MP)

# need data set of unique short cite by expt_num
# in order to calculate total number of infants

# need data set of unique short cite by condition

sum_dat <- dat[!duplicated(dat[c("short_cite", "same_infant")]),]
time_wind_dat <- dat[!duplicated(dat[c("short_cite", "offset", "post_nam_dur")]),]
distract_dat <- dat[!duplicated(dat[c("short_cite", "object_pair")]),]
mix_co_mp <- dat[!duplicated(dat[c("short_cite", "word_correct_and_MP")]),]
vocab_dat <- dat[!duplicated(dat[c("short_cite")]),]

```

# Systematic Study Collection

The database contains data from `r length(unique(dat$short_cite))` papers consisting of data from `r sum(sum_dat$n_1)` infants.

The next table shows what type of publications were included in our meta-analysis
```{r DescriptivesPubtype, echo = FALSE}

pub_info <- dat %>%
  group_by(publication_status) %>%
  summarize(n_unique = n_distinct(short_cite))

kable(pub_info)

```

# Effect Size Calculation

## Type of Dependent Variable

The table below summarized the type of dependent variable and subsequent comparison reported in each paper.

```{r DescriptivesComparison, echo = FALSE}
db_ET_correct$within_measure_descriptive <- ifelse(db_ET_correct$within_measure == "post",
 "post-naming phase compared with chance (=50%)",
ifelse(db_ET_correct$within_measure == "pre_post", "post-naming compared to pre-naming phase", "post-pre difference score compared with chance (=0)"))

comparison_info <- db_ET_correct %>%
  group_by(within_measure_descriptive) %>%
  summarize(n_unique = n_distinct(short_cite))

kable(comparison_info)
#table(db_ET_MP$es_method)
```

## Data Used to Calculate Effect Sizes

The table below shows based on which data we calculated effect sizes. If both raw means/standard deviations as well as t-values were available, we used raw means/standard deviations to calculate effect sizes.

```{r DescriptivesES, echo = FALSE}
#how did we calculate effect sizes?

es_info <- dat %>%
  group_by(es_method) %>%
  summarize(n_unique = n_distinct(short_cite))

kable(es_info)
#table(dat$es_method, dat$short_cite)
# only missing from:
# Bailey & Plunkett (2002) <- crazy old data
# Renner (2017) <- just finished phd
# Tamasi (2016) <- standard deviations are wrong
#table(db_ET_MP$es_method)

```

# Meta-Analysis

## Object Identification

### Correct words

Meta-analytic effect for correctly pronounced words in object identification.

```{r MAcorrect}

rma_correct = rma.mv(g_calc, g_var_calc, data = db_ET_correct, random = ~ collapse | short_cite)

summary(rma_correct)

```

### Mispronounced words

Meta-analytic effect for mispronounced words in object identification.

```{r MAMP}

rma_MP = rma.mv(g_calc, g_var_calc, data = db_ET_MP, random = ~ collapse | short_cite)

summary(rma_MP)

```

## Mispronunciation Sensitivity

Meta-analytic effect for mispronunciation senstivity.

```{r MPEffect}
db_ET_correct$condition <- 1
db_ET_MP$condition <- 0

dat <- bind_rows(db_ET_correct, db_ET_MP)

rma_MPeffect <- rma.mv(g_calc, g_var_calc, mods = ~condition, data = dat, random = ~ collapse | short_cite)

summary(rma_MPeffect)

```

## Age Moderator Effects

### Object identification correct words

Age moderator effects for object identification for correctly pronounced words.

```{r MAcorrect_age}
rma_correct_age = rma.mv(g_calc, g_var_calc, mods = ~age.C, data = db_ET_correct, random = ~ collapse | short_cite)

summary(rma_correct_age)

```

### Object identification mispronounced words

Age moderator effects for object identification for mispronounced words.

```{r MAMP_age}

rma_MP_age = rma.mv(g_calc, g_var_calc, mods = ~age.C, data = db_ET_MP, random = ~ collapse | short_cite)

summary(rma_MP_age)
```

### Mispronunciation sensitivity

Age moderator effects for mispronunciation sensitivity

```{r MPEffect_age}
db_ET_correct$condition <- 1
db_ET_MP$condition <- 0

dat <- bind_rows(db_ET_correct, db_ET_MP)

rma_MPeffect_age <- rma.mv(g_calc, g_var_calc, mods = ~age.C*condition, data = dat, random = ~ collapse | short_cite)

summary(rma_MPeffect_age)

```

# Plotting Mispronunciation Sensitivity by Age

This is the plot used in Figure 2. Effect sizes for correct pronunciations (yellow) and mispronunciations (blue) by participant age. Point size is inverse variance. The dashed line indicates zero.

```{r FunnelPrep, echo = FALSE}

#Themes and plot
apatheme=theme_bw()+
  theme(#panel.grid.major=element_blank(),
    #panel.grid.minor=element_blank(),
    #panel.border=element_blank(),
    axis.line=element_line(),
    text=element_text(family='Times', size=25),
    legend.position='none')

```

```{r PlotMPEffect, echo = FALSE}

dat$condition_label = ifelse(dat$condition==1, "Correct", "Mispronunciation")

# Color Blind palette:
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

p <- ggplot(dat, aes(mean_age_1/30.44, g_calc, color = condition_label)) + 
  geom_point(aes(size = weights_g),show.legend=FALSE) + 
  geom_line(y= 0, linetype="dotted") + 
  geom_smooth(method = "lm", formula = y ~ log(x), aes(weight=weights_g)) +
  scale_colour_manual(values=cbPalette)+
  apatheme +
  theme(legend.title = element_blank(),
        legend.position = "bottom") + 
  xlab("Age in months") + 
  ylab("Effect size Hedges' g")

#p

#min(dat$mean_age_1/30.44)
#max(dat$mean_age_1/30.44)

#ggsave("figures/AgeEffect_log.jpg", p,height= 7,width= 6)

  #jpeg(filename = "figures/AgeEffect_log.jpg", 
 #      width = 600, height = 400, units = "px")

  p

  #dev.off()

```

# Supplementary Analyses for Discussion

This section includes supplementary analyses discussed in the Discussion section

## Vocabulary Data Available

This table summarizes which papers also measured vocabulary in the infants they tested.

```{r VocabularyData}

vocab_info <- vocab_dat %>%
  mutate(has_vocab = ifelse(!is.na(r_comprehension) & !is.na(r_production), "comprehension & production", 
                            ifelse(!is.na(r_comprehension), "just comprehension",
                            ifelse(!is.na(r_production), "just production", "none")))) %>%
  group_by(has_vocab) %>%
  summarize(count = n())

kable(vocab_info)

```

## Mispronunciation Sensitivity at Youngest Ages

Mispronunciation sensitivity in the papers in our dataset with the youngest ages (< 12-months-of-age; Bergelson & Swingley, 2017; Mani & Plunkett, 2007; Zesiger et al., 2012).

```{r YoungAge, echo = FALSE}
youngest <- subset(dat, min_age_1 < 365)

#unique(youngest$short_cite)
youngest$cond_lab <- ifelse(youngest$condition == 1, "correct", "misp")
p <- ggplot(youngest, aes(x=cond_lab, y = g_calc, color = cond_lab)) + geom_boxplot()

p

```

## Known vs Novel Distractor Images

This table summarizes which papers used a known or novel distractor image.

```{r DescriptivesDISTRACT, echo = FALSE}
#What kind of distractor was used?

distr_info <- distract_dat %>%
  group_by(object_pair) %>%
  summarize(count = n())

kable(distr_info)

```

### Does distractor familiarity (known vs. novel) impact mispronunciation sensitivity?

Preliminary moderator test examining the influence of distractor familiarity on mispronunciation sensitivity

```{r DistractorTypeAge}

rma_Distractor <- rma.mv(g_calc, g_var_calc, mods = ~condition*as.factor(object_pair), data = dat, random = ~ collapse | short_cite)

summary(rma_Distractor)

```

## Plotting Influence of Distractor Familiarity

```{r PlotDistFamEffect_NoAge, echo = FALSE}

dat$condition_label = ifelse(dat$condition==1, "correct", "mispronunciation")
dat$dist_code <- ifelse(dat$object_pair == "familiar_familiar", "known", "novel")

p <- ggplot(dat, aes(dist_code, g_calc, color = condition_label)) + 
  geom_boxplot() + 
  #geom_smooth(method = "lm", formula = y ~ log(x), aes(weight=weights_g)) + 
  scale_color_manual(values=wes_palette(name="Darjeeling"))+
  theme_bw()+
  theme(text = element_text(size=25),
        legend.title = element_blank(),
        legend.position = "bottom") + 
  xlab("Distractor Familiarity") + 
  geom_hline(yintercept = 0, linetype="dotted") + 
  ylab("Effect size Hedges' g")
p

#ggsave("figures/Distractor_fam_log.jpg", p)

```