### 2018-05-09 CogSci_Publication_bias.Rmd
updated code for response to CogSci reviewers
- changed color to dark blue for age by effect size plot
- changed axis and overall titles for funnel plot

---
title: "Mispronunciation Sensitivity Publication Bias for CogSci"
author: "Christina Bergmann and Katie Von Holzen"
output: 
  pdf_document:
    toc: true
---
```{r Setup, echo = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, tidy = TRUE, message = FALSE, error = TRUE)

### Load libraries
library(metafor)
library(pwr)
library(ggplot2)
library(grid)
library(gridExtra)

```

# Preparation
Read in data and tidy up dataset

```{r ReadIn,  echo = FALSE}
#Get the data
source("scripts/Meta_Analysis.R")

# change name of mispronunciation sensitivity object
db_mpe <- dat
rm(dat)

```

# Funnel Plot Asymmetry
Use ranktest() to check for funnel plot asymmetry. Both tests show evidence of significant asymmetry.

## Object Identification Correct Words

```{r FunnelPlotAsymm_correct, echo = FALSE}

# Publication bias can become visible in funnel plot asymmetry. 
# Metafor comes with several options to check this, I chose ranktest and regtest
# correct object identification
ranktest(rma_correct)

```

## Object Identification Mispronounced Words

```{r FunnelPlotAsymm_misp, echo = FALSE}

# mispronounced object identification
ranktest(rma_MP)

```

# Draw Funnel Plots

### Plots adapted from Sakaluk, 2016, see also Black & Bergmann, 2017
### https://sakaluk.wordpress.com/2016/02/16/7-make-it-pretty-plots-for-meta-analysis/

```{r FunnelPrep, echo = FALSE}

#Themes and plot
apatheme=theme_bw()+
  theme(#panel.grid.major=element_blank(),
    #panel.grid.minor=element_blank(),
    #panel.border=element_blank(),
    axis.line=element_line(),
    text=element_text(family='Times', size=25),
    legend.position='none')

```

## Funnel Plot for Correct Object Identification

```{r FunnelCorrect, echo = FALSE}

#Rename a bunch of things for ease
dat_co = db_ET_correct
#dat_co=db_ET_correct
dat_co <-
  dat_co %>% 
  rename(cite = short_cite,
         yi = g_calc,
         vi = g_var_calc)

#Reorder bibliographic info based on value of g (yi), so effect sizes can be plotted in descending order

dat_co <-
  dat_co %>% 
  select(cite, expt_num, same_infant, mean_age_1, yi, vi) %>% 
  mutate(study_ref = paste(cite, expt_num, same_infant, sep=',')) %>% 
  arrange(desc(yi))

#Get standard errors from variances
dat_co$se = sqrt(dat_co$vi)

#Calculate 95% CI values
dat_co$lowerci = (-1.96*dat_co$se)+dat_co$yi
dat_co$upperci = (1.96*dat_co$se)+dat_co$yi

#Store the meta-analytic estimate and its standard error from whatever model you run
#summary(rma_correct)
estimate_co = coef(summary(rma_correct))$estimate
se_co = coef(summary(rma_correct))$se

#Store a vector of values that spans the range from 0
#to the max value of impression (standard error) in your dataset.
#Make the increment (the final value) small enough (I choose 0.001)
#to ensure your whole range of data is captured
se.seq_co=seq(0, max(dat_co$se), 0.001)

#Now, compute vectors of the lower-limit and upper limit values for
#the 95% CI region, using the range of SE that you generated in the previous step, 
#and the stored value of your meta-analytic estimate.
ll95_co = estimate_co-(1.96*se.seq_co)
ul95_co = estimate_co+(1.96*se.seq_co)

#You can do this for a 99% CI region too
ll99_co = estimate_co-(3.29*se.seq_co)
ul99_co = estimate_co+(3.29*se.seq_co)

#And finally, do the same thing except now calculating the confidence interval
#for your meta-analytic estimate based on the stored value of its standard error
meanll95_co = estimate_co-(1.96*se_co)
meanul95_co = estimate_co+(1.96*se_co)

#Now, smash all of those calculated values into one data frame (called 'dfCI').
#You might get a warning about '...row names were found from a short variable...'
#You can ignore it.
dfCI_co = data.frame(ll95_co, ul95_co, ll99_co,
                  ul99_co, se.seq_co, estimate_co, meanll95_co, meanul95_co)

#Now we can actually make the funnel plot.
#Using your original data-frame, map standard error to your x-axis (for now) and Zr to your y-axis
fp_co = ggplot(aes(x = se, y = yi), data = dat_co) +
  #Regression line for the FP asymmetry
  geom_smooth(aes(x = se, y = yi), method = "lm", colour = "darkgrey", alpha = .5, se = FALSE, data = dat_co) +
  #Add your data-points to the scatterplot
  geom_point(size = 2.5, colour="black") +
  #
  #Give the x- and y- axes informative labels
  xlab('Standard Error') + ylab('Hedge\'s g')+
  # give it a title
    ggtitle("Correct") +
    # make sure both plots have same scale
    # still ylim because we haven't flipped yet
    #coord_cartesian(ylim = c(0.5, 0))+
  #Now using the 'dfCI' data-frame we created, plot dotted lines corresponding
  #to the lower and upper limits of your 95% CI region
  #And dashed lines corresponding to your 99% CI region
  #Add lines corresponding to 0 and estimate
  geom_line(aes(x = se.seq_co, y = 0), linetype = 'solid', data = dfCI_co) +
  geom_line(aes(x = se.seq_co, y = estimate_co), linetype = 'dashed', data = dfCI_co) +
  geom_line(aes(x = se.seq_co, y = ll95_co), linetype = 'dotted', data = dfCI_co) +
  geom_line(aes(x = se.seq_co, y = ul95_co), linetype = 'dotted', data = dfCI_co) +
  #  geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
  #  geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
  #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
  #geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI) +
  #geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI) +
  #Reverse the x-axis ordering (se) so that the tip of the funnel will appear
  #at the top of the figure once we swap the x- and y-axes...
  scale_x_reverse()+
  #Specify the range and interval for the tick-marks of the y-axis (Zr);
  #Choose values that work for you based on your data
  #scale_y_continuous(breaks=seq(-.45,0.8,0.25))+
  #And now we flip the axes so that SE is on y- and Zr is on x-
  coord_flip()+
  #Finally, apply my APA-format theme (see code at end of post).
  #You could, alternatively, specify theme_bw() instead.
  apatheme

#Call the pretty funnel plot
fp_co
#ggsave("figures/FunnelPlot_correct.pdf")

```

## Funnel Plot for Mispronunciation Object Identification

```{r FunnelMisp, echo = FALSE}

  #Rename a bunch of things for ease
dat_mp = db_ET_MP  
#dat_mp=db_ET_MP
  dat_mp <-
    dat_mp %>% 
    rename(cite = short_cite,
           yi = g_calc,
           vi = g_var_calc)

  #Reorder bibliographic info based on value of g (yi), so effect sizes can be plotted in descending order

  dat_mp <-
    dat_mp %>% 
    select(cite, expt_num, same_infant, mean_age_1, yi, vi) %>% 
    mutate(study_ref = paste(cite, expt_num, same_infant, sep=',')) %>% 
    arrange(desc(yi))

  #Get standard errors from variances
  dat_mp$se = sqrt(dat_mp$vi)

  #Calculate 95% CI values
  dat_mp$lowerci = (-1.96*dat_mp$se)+dat_mp$yi
  dat_mp$upperci = (1.96*dat_mp$se)+dat_mp$yi

  #Store the meta-analytic estimate and its standard error from whatever model you run (substitute your own values)
  #summary(rma_MP)
  estimate_mp = coef(summary(rma_MP))$estimate
  se_mp = coef(summary(rma_MP))$se

  #Store a vector of values that spans the range from 0
  #to the max value of impression (standard error) in your dataset.
  #Make the increment (the final value) small enough (I choose 0.001)
  #to ensure your whole range of data is captured
  se.seq_mp=seq(0, max(dat_mp$se), 0.001)

  #Now, compute vectors of the lower-limit and upper limit values for
  #the 95% CI region, using the range of SE that you generated in the previous step, 
  #and the stored value of your meta-analytic estimate.
  ll95_mp = estimate_mp-(1.96*se.seq_mp)
  ul95_mp = estimate_mp+(1.96*se.seq_mp)

  #You can do this for a 99% CI region too
  ll99_mp = estimate_mp-(3.29*se.seq_mp)
  ul99_mp = estimate_mp+(3.29*se.seq_mp)

  #And finally, do the same thing except now calculating the confidence interval
  #for your meta-analytic estimate based on the stored value of its standard error
  meanll95_mp = estimate_mp-(1.96*se_mp)
  meanul95_mp = estimate_mp+(1.96*se_mp)

  #Now, smash all of those calculated values into one data frame (called 'dfCI').
  #You might get a warning about '...row names were found from a short variable...'
  #You can ignore it.
  dfCI_mp = data.frame(ll95_mp, ul95_mp, ll99_mp,
                       ul99_mp, se.seq_mp, estimate_mp, meanll95_mp, meanul95_mp)

  #Now we can actually make the funnel plot.
  #Using your original data-frame, map standard error to your x-axis (for now) and Zr to your y-axis
  fp_mp = ggplot(aes(x = se, y = yi), data = dat_mp) +
    #Regression line for the FP asymmetry
    geom_smooth(aes(x = se, y = yi), method = "lm", colour = "darkgrey", alpha = .5, se = FALSE, data = dat_mp) +
    #Add your data-points to the scatterplot
    geom_point(size = 2.5, colour="black") +
    #
  #Give the x- and y- axes informative labels
  xlab('Standard Error') + 
    # remove y label because it will be combined with the other funnel plot
    ylab('Hedge\'s g')+
    # give it a title
    ggtitle("Mispronunciation") +
    # make sure both plots have same scale
    # still xlim because we haven't flipped yet
    #coord_cartesian(xlim = c(0.5, 1))+
    #Now using the 'dfCI' data-frame we created, plot dotted lines corresponding
    #to the lower and upper limits of your 95% CI region
    #And dashed lines corresponding to your 99% CI region
    #Add lines corresponding to 0 and estimate
    geom_line(aes(x = se.seq_mp, y = 0), linetype = 'solid', data = dfCI_mp) +
    geom_line(aes(x = se.seq_mp, y = estimate_mp), linetype = 'dashed', data = dfCI_mp) +
    geom_line(aes(x = se.seq_mp, y = ll95_mp), linetype = 'dotted', data = dfCI_mp) +
    geom_line(aes(x = se.seq_mp, y = ul95_mp), linetype = 'dotted', data = dfCI_mp) +
    #  geom_line(aes(x = se.seq, y = ll99), linetype = 'dashed', data = dfCI) +
    #  geom_line(aes(x = se.seq, y = ul99), linetype = 'dashed', data = dfCI) +
    #Now plot dotted lines corresponding to the 95% CI of your meta-analytic estimate
    #geom_segment(aes(x = min(se.seq), y = meanll95, xend = max(se.seq), yend = meanll95), linetype='dotted', data=dfCI) +
    #geom_segment(aes(x = min(se.seq), y = meanul95, xend = max(se.seq), yend = meanul95), linetype='dotted', data=dfCI) +
    #Reverse the x-axis ordering (se) so that the tip of the funnel will appear
    #at the top of the figure once we swap the x- and y-axes...
    scale_x_reverse()+
    #Specify the range and interval for the tick-marks of the y-axis (Zr);
    #Choose values that work for you based on your data
    #scale_y_mpntinuous(breaks=seq(-.45,0.8,0.25))+
    #And now we flip the axes so that SE is on y- and Zr is on x-
    coord_flip()+
    #Finally, apply my APA-format theme (see previous).
    #You could, alternatively, specify theme_bw() instead.
    apatheme +
    theme(axis.title.y = element_blank())

  #Call the pretty funnel plot
  fp_mp
#  ggsave("figures/FunnelPlot_misp.pdf")

```

## Combine both funnel plots for figure

Below is the plot used in Figure 1. Funnel plots for object identification for correct (left) and mispronounced (right) words, plotting the standard error of the effect size in relation to effect size. The black line marks zero, the dashed grey line marks the effect estimate, and the grey line marks funnel plot asymmetry.

```{r FunnelCombo, echo = FALSE}

  jpeg(filename = "figures/Funnel_Plots_Object_Identification.jpg", 
       width = 600, height = 400, units = "px")

  p <-   grid.arrange(fp_co, fp_mp, nrow = 1)
                      #top = textGrob("Object Identification", gp=gpar(fontsize=25)))
                      #top = "Object Identification")

  dev.off()

    grid.draw(p)

```

# P-curve

## Export Data for P-curve

This step creates a text file that can be used in the p-curve app:
http://www.p-curve.com/app4/

for correct word object identification see "p_curve_co.txt" file

for mispronounced word object identification see "p_curve_mp.txt" file

Just copy and paste the .txt onto the website and voila!

```{r PcurveExport, echo = FALSE}

textfile_co = "p_curve_app/p_curve_co.txt"

# make sure the file is made new with our loop
if (file.exists(textfile_co)) {file.remove(textfile_co)}

for(line in 1:length(db_ET_correct$n_1)){
  if(!is.na(db_ET_correct[line,]$t)){
    newline = paste("t(", db_ET_correct[line,]$n_1-1, ")=",db_ET_correct[line,]$t, sep = "")
    write(newline, file = textfile_co, append = TRUE)
    #cat("\n")
    #write(newline, file = textfile, append = TRUE)    
  }
}

textfile_mp = "p_curve_app/p_curve_mp.txt"

# make sure the file is made new with our loop
if (file.exists(textfile_mp)) {file.remove(textfile_mp)}

for(line in 1:length(db_ET_MP$n_1)){
  if(!is.na(db_ET_MP[line,]$t)){
    newline = paste("t(", db_ET_MP[line,]$n_1-1, ")=",db_ET_MP[line,]$t, sep = "")
    write(newline, file = textfile_mp, append = TRUE)
    #cat("\n")
    #write(newline, file = textfile, append = TRUE)    
  }
}

```

# Calculate Power

```{r Power_exist_descript, echo = FALSE}
d = rma.mv(g_calc, g_var_calc, data = db_mpe, random = ~ 1 | short_cite)$b[,1]

# estimate from mispronunciation sensitivity
mp_est <- round(coef(summary(rma_MPeffect))[2,]$estimate, 2)

# one-sided t-test
ptt <- pwr.t.test(d = mp_est, sig.level = .05, type = "paired", power = .80, alternative = "greater")

```

The median sample size across studies in our dataset was `r median(db_mpe$n_1, na.rm=TRUE)`. To study infants' mispronunciation senstivity, which we estimate to have an effect size of `r mp_est`, this test gives the number of participants (n = `r round(ptt$n, 2)`) required in a within-participant design using a one-sided t-test to conduct a study which has 80% power.

```{r Power_plan, echo = FALSE}

# one-sided t-test
pwr.t.test(d = mp_est, sig.level = .05, type = "paired", power = .80, alternative = "greater")

```