### 2020-02-07 Untitled/Untitled.Rmd
Dev Psych submission

---
title: "Template Title"
author: "Your Name"
date: "The Date"
output:
  bookdown::pdf_document2:
    number_sections: no
    toc: no
    latex_engine: xelatex
header-includes:    
 - \usepackage{tipa}    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If I knit an Rmd document that includes UTF-8 characters outside the ASCII range (for instance, phonetic symbols like \textipa{[E]} or \textipa{[O]}), they never survive the transition to PDF (with output: `papaja::apa6_pdf`) or to .docx (with output: `papaja::apa6_word`).